### 2020-05-11 thesis_template/0_Prelim/abs.tex
updated Abstract

Many existing dataflow tools use Java Virtual Machine (JVM). 
Memory strategies in JVM, such as Garbage Collection (GC), may lead to significant overhead in Big Data processing. 
Apache Spark and Apache Flink use complex objects to manipulate and transfer a tremendous amount of data. 
Generating many of these complex objects in memory forces GC to rearrange the objects in memory frequently waisting computation.

Considered problems in memory management in JVM, developing Big Data processing tools with system languages can be the solution.
By using a system language, a developer has control on the memory management. Therefore, one can implement systems with more optimized memory management strategies.
We select Rust as a good candidate for the development of Big Data processing tools, due to its ability to write memory-safe and fearless concurrent codes with its concept of ownership.

There may be many possible strategies to optimize memory management for Big Data processing in Rust programming: 
selection of different variable types, use of Reference Counting, and multithreading with Atomic Reference Counting.
We conduct several experiments to assess how much these different memory management strategies differ runtime performance.

Our experiments focus complex object manipulation and common Big Data processing with various memory management.
The results show significant different runtime hits among these different memory strategies.

>>>>

Many existing modern cloud-based data processing systems such as Hadoop, Spark or Flink use
Java Virtual Machine (JVM) and taking full advantage of features such as automated memory management in JVM
including Garbage Collection (GC) which may lead to a significant overhead.
Dataflow-based systems like Spark allow programmers to define complex objects in a
host language like Java to manipulate and transfer tremendous amount of data.

% Generating many of these complex objects in memory forces GC to rearrange the objects in memory frequently waisting computation.
% Considered performance overhead in memory management in JVM makes system languages to be a better choice to
% develop systems for Big Data processing.

System languages like C++ or Rust seems to be a better choice to
develop systems for Big Data processing becasuse they do not relay on JVM.
By using a system language, a developer has full control on the memory management.
% Therefore, one can implement systems with more optimized memory management strategies.
We found Rust programming language to be a good candidate due to its ability to write memory-safe and
fearless concurrent codes with its concept of memory ownership and borrowing.
Rust programming language includes many possible strategies to optimize memory management for Big Data processing including
selection of different variable types, use of Reference Counting, and multithreading with Atomic Reference Counting.

In this thesis, we conducted an experimental study to assess how much these different memory management
strategies differ regarding overall runtime performance.
Our experiments focusses on complex object manipulation and common Big Data processing patterns
with various memory management. Our experimental results indicate a significant
difference among these different memory strategies regarding data processing performance.
